---
title: "Text Analysis"
format: html
author: "Morgan Ryan"
editor_options: 
  chunk_output_type: console
resource_files:
- "Text-Analysis.html"
---

```{r}
#| include: FALSE

#libraries and data

library(tidyverse)
library(tidytext)
library(textdata)
library(wordcloud)
library(wordcloud2)
library(viridis)
library(ggthemes)
library(gutenbergr)
library(pixarfilms)
library(RColorBrewer)

spot_mill <- read_csv("~/SDS264_F24/SDS264/Data/spotify_millsongdata.csv")
bing_sentiments <- get_sentiments(lexicon = "bing")

```

Description:
I wanted to do text analysis on an artist I frequently listen to. I love the emotion in fleetwood mac songs and wanted to see what I could uncover. As soon as I began digging into the data, I noticed words like ah, ooh, and more were sneaking their way into my graphs, tables, and even word clouds. To fix that, I immediately filtered them out, then I also chose to filter out the apostrophes. 

What did I find? One of my favorite finds was that the most commonly used word in her songs was love. This feels fitting knowing Stevie Nicks history with Lindsey Buckingham. Moving on to a side by side comparision of the top twenty words divided up into positive and negative categories, I wasn't surprised to find more negatives then positives. While Fleetwood Mac's music is catchy, I wouldn't define it as overwhelmingly positive or happy. I thought there was a creative spread of negative words like hard, blow, bad, fool, wrong, lonely. They tell a story, we know there was a breakup between Nicks and Buckingham and I feel as though their relationship was represented in here by the sheer number of angry words used. In the first word cloud, we can see how love is the largest, surrounded by lonely, wrong, free, fool, win, etc. This is an emotional word cloud and I would expect nothing less. Curious about diving deeper into the negative feelings, I made a word cloud based on words that correlated with anger. I was not surprised that Bad was the first word but it was surrounded by lonely, feeling, lose, crazy, shatter, etc. Seeing the data laid out demonstrated a clear picture of how Stevie Nicks incorporated her anger into song. Being inspired by my anger word cloud, I created another graph that would show me which songs specifically were the most "angry". I hope to listen to them with this in mind and see if I agree. Overall, this text analysis allowed me to look further into one of my favorite bands and supported my digging into the band relationships and seeing which songs correlated with specific anger sentiments. When I do more research into them, I hope to keep these findings in mind. If I wanted to continue investigating, I might try to add data with a release date to each song and see how the anger sentiments change over time.


First, finding the data I want to work with.

```{r}
fleet_mac <- spot_mill |>
  filter(str_detect(artist, "^(f|F).*c$")) |> #Str function, regex #1
  filter(artist == "Fleetwood Mac") |>
  select(-link, -artist)

fleet_mac
```

Next, tokenize and and tidy the data 

```{r}
tidy_fleet_mac <- fleet_mac |>
  mutate(line = row_number()) |>
  unnest_tokens(word, text, token = "words")

tidy_fleet_mac

smart_stopwords <- get_stopwords(source = "smart")
```

*Table of expressive words used in songs*
- expressive = oohs, ahs, etc
I found this partially on accident after working on removing them from my text analysis, but thought it would be fun to include.  

```{r}
tidy_fleet_mac |>
  select(-line) |>
  mutate(expression = str_count(word, "([^d]((oo|a+|yea)h)|who+a)")) |> #str_function, regex #2
  filter(expression > 0) |>
  count(word, expression, sort = TRUE)

```

*Graph of most common 20 words*
I'm keeping my eye on words that might represent anger, or hint at any tensions between Nicks and Buckingham. We can see how love is prominently first, along with a few other positive words. Before words like women and she enter the charts, this is interesting as maybe she's referencing Buckingham with another women.

```{r}
tidy_fleet_mac |>
  anti_join(smart_stopwords) |>
  count(word, sort = TRUE) |>
  filter(word != "NA", str_detect(word, "[^((oo|a+|yea)h)|who+a]")) |> #Filter out expressive words
  mutate(word = str_replace_all(word, "'", ""))  |> #str_function, regex #3, removing ' in words
  slice_max(n, n = 20) |>
  ggplot(aes(fct_reorder(word, n), n, fill = n)) +
  geom_col() +
  coord_flip() +
  theme_clean() +
  labs(title = "Most Common Words in Fleetwood Mac Songs",
       x = "Word",
       y = "Count") 

```

*Table and Graph of top 10 positive and negative words*
I'm not surprised at the higher count of negative words. I feel like Stevie Nicks also weaponizes the "positive" words to sometimes be used demeaningly and often with much irony. As the themes in her songs are often of heartbreak and have a deep emotional depth. 

```{r}
fleet_mac_sentiments <- tidy_fleet_mac |>
  inner_join(bing_sentiments) |>
  count(sentiment, word, sort = TRUE) |>
  group_by(sentiment) |>
  slice_max(n, n = 10) |>
  ungroup()

fleet_mac_sentiments

fleet_mac_sentiments |>
  #filtering to get rid of ooh's/yeah/etc
  filter(str_detect(word, "[^((oo|a+|yea)h)|who+a]")) |>
  ggplot(aes(x = fct_reorder(word, n), y = n, fill = sentiment)) +
    geom_col(show.legend = FALSE) +  
    coord_flip() +
    facet_wrap(~ sentiment, scales = "free") +
  theme_clean() +
  labs(title = "Top 10 Most Common Positive/Negative Sentiments in Fleetwood Mac Songs",
       x = "Count",
       y = "Word")

```

*Basic Word Cloud Using Bing Sentiments*
Sticking with the irony theme, Love being the largest here really emphasizes that.
```{r}
#| warning: FALSE
#| message: FALSE

words_fleet_mac <- tidy_fleet_mac |>
  anti_join(stop_words, by = join_by(word)) |>
  inner_join(bing_sentiments) |>
  count(sentiment, word, sort = TRUE) |>
  group_by(sentiment) |>
  filter(str_detect(word, "[^((oo|a+|yea)h)|who+a]")) |>
  arrange(desc(n)) |>
  mutate(colors = ifelse(sentiment == "positive", "lightgreen", "tomato"))

wordcloud(
  words = words_fleet_mac$word, 
  freq = words_fleet_mac$n, 
  max.words = 100, 
  random.order = FALSE,
  color = words_fleet_mac$colors
)
```

*Word Cloud Using NCR Sentiments - Anger*
To try and get around the ironic uses of positive words, I decided to focus on anger. I feel like the selection of negative words she uses are much more direct and tell a more meaningful story. Like honest, broken, lose, crazy, blame, all contribute to a complex story and relationships that she has.
```{r}
ncr_sentiments <- get_sentiments(lexicon = "nrc")

angry_fleet_mac <- ncr_sentiments |>
  filter(sentiment == "anger") |>
  inner_join(tidy_fleet_mac) |>
  anti_join(stop_words, by = join_by(word)) |>
  count(word) |>
  filter(word != "NA") |>
  arrange(desc(n))

wordcloud(
  words = angry_fleet_mac$word, 
  freq = angry_fleet_mac$n, 
  max.words = 100, 
  random.order = FALSE,
  color = brewer.pal(6, "Spectral")
)
```

*Graph as another way to visualize anger word count using song*
Here I wanted to explore these negative words with the song as context. Now, when compared to a list I found online, none of these songs would be listed as the top "angry" songs by Fleetwood Mac. But I hope to listen to them and see if they meaningfully add to the narrative.

```{r}
ncr_sentiments |>
  filter(sentiment == "anger") |>
  inner_join(tidy_fleet_mac) |>
  anti_join(stop_words, by = join_by(word)) |>
  count(song) |>
  slice_max(n, n = 10) |>
  arrange(desc(n)) |>
  ggplot(aes(x = fct_reorder(song, n), y = n, fill = song)) +
    geom_col(show.legend = FALSE) +  
    coord_flip() +
    theme_clean() +
    labs(
      title = "Top Ten Number of Words Associated with Anger by Fleetwood Mac Song",
      x = "Song",
      y = "Anger Word Count")

```

*A Note About References to Stevie Nick's Song Writing*
It should be emphasized that while I reference Nicks as writing the songs, that Buckingham also wrote songs as song are also from his point of view. This doesn't necessarily alter the narrative too much as they both conveyed complex emotions through their music. In a separate project, I might be able to assign who wrote what song and see hwo that changes what I've found.

